---
title: "MaxQuant to MSstats preprocessing Report"
author: "Miguel Cosenza"
date: "28 11 2019"
output: 
  html_document:
    keep_md: true
---

```{r, echo = FALSE, include=FALSE}
knitr::opts_chunk$set(
  fig.path = "Figures_Sample_Report_2/",
  echo = TRUE
)
```

# MaxQuant to MSstats formating, normalization, simple Quality control and generation of wide-formatted tables (for limma input)  

## Load required packages

```{r load_packages}
library(tidyverse)
library(MSstats)
```

## Load input data 

```{r load_input_data}
evidence <- read.table(file = here::here("evidence.txt"),
                       sep = "\t",
                       header = TRUE)

annotation <- read.csv(file = here::here("annotation.csv"),
                       header = TRUE)

proteingroups <- read.table(file = here::here("proteinGroups.txt"),
                            sep = "\t",
                            header = TRUE)
```

## Transformation from Max Quant format to MSstats format 

With all parameters set to default and including proteins identified with just 1 peptide.  

```{r mqtomstformat1}
if(dir.exists(here::here("MSstats_Output_data")) == FALSE){dir.create(here::here("MSstats_Output_data"))}

if(file.exists(x = here::here("MSstats_Output_data/msts_data_w1pep.Rda")) == FALSE){ # check if Rda file with the data is already available. 
      
      msts_data_w1pep <- MaxQtoMSstatsFormat(evidence = evidence,
                                         annotation = annotation,
                                         proteinGroups = proteingroups,
                                         removeProtein_with1Peptide = FALSE)
      
      save(msts_data_w1pep, file = here::here("MSstats_Output_data/msts_data_w1pep.Rda")) # save Rda file for faster loading later if required
      
      if(dir.exists(here::here("MSstats_Output_data/MSstats_formated_tables")) == FALSE){dir.create(here::here("MSstats_Output_data/MSstats_formated_tables"))}
      
      write.csv(x = msts_data_w1pep, file = here::here("MSstats_Output_data/MSstats_formated_tables/msstats_data_protsw1pep.csv")) # protsw1pep means 'this file contains proteins that were identified only with 1 peptide". 
} else {
      
      load(here::here("MSstats_Output_data/msts_data_w1pep.Rda")) # load the Rda file if available.
      
}
```

## Correction of protein IDs  

Get the first listed protein ID when many Uniprot IDs are listed into one Protein Group (many identified peptides matching several proteins in the database)

```{r correct_proteinIDs}
msts_formated_data <- dplyr::mutate(msts_data_w1pep,
                             ProteinName = stringr::str_remove_all(ProteinName, ";.*$")) %>%
                      dplyr::mutate(ProteinName = stringr::str_trim(ProteinName))
```

## Normalization 

Using default parameters (log2 transformation, normalization by `equalizeMedians` and using all features for quantification).  

```{r dataprocess1}
dp_all_log2_eqmNorm <- dataProcess(msts_formated_data,
                                   logTrans=2,
                                   normalization="equalizeMedians",
                                   nameStandards=NULL,
                                   address="",
                                   fillIncompleteRows=TRUE,
                                   featureSubset="all",
                                   remove_uninformative_feature_outlier=FALSE,
                                   n_top_feature=3,
                                   summaryMethod="TMP",
                                   equalFeatureVar=TRUE,
                                   censoredInt="NA",
                                   cutoffCensored="minFeature",
                                   MBimpute=TRUE,
                                   remove50missing=FALSE,
                                   maxQuantileforCensored=0.999,
                                   clusters=NULL)
```

```{r dataprocess2, include=FALSE}
dp_all_log2_noNorm <- dataProcess(msts_formated_data,
                                   logTrans=2,
                                   normalization=FALSE,
                                   nameStandards=NULL,
                                   address="",
                                   fillIncompleteRows=TRUE,
                                   featureSubset="all",
                                   remove_uninformative_feature_outlier=FALSE,
                                   n_top_feature=3,
                                   summaryMethod="TMP",
                                   equalFeatureVar=TRUE,
                                   censoredInt="NA",
                                   cutoffCensored="minFeature",
                                   MBimpute=TRUE,
                                   remove50missing=FALSE,
                                   maxQuantileforCensored=0.999,
                                   clusters=NULL)
```

## Quality control plots  

#### Normalized data
```{r qcplot1, fig.width=6.5, fig.height=4}
dataProcessPlots(data = dp_all_log2_eqmNorm, 
                 type="QCplot", 
                 width=7, height=7,
                 which.Protein = 'allonly',
                 address=FALSE)
```

#### Not normalized for comparison  

```{r qcplot2, fig.width=6.5, fig.height=4}
dataProcessPlots(data = dp_all_log2_noNorm, 
                 type="QCplot", 
                 width=7, height=7,
                 which.Protein = 'allonly',
                 address=FALSE)
```


## Getting tabular data into wide format as an input for Limma  

The `RunlevelData` section of the `list` generated after `dataProcess` by MSstats contains the tabular data.  
The 'Run level data' has information of LogIntensities for proteins per MS Run. The first step before transforming the data into a wide format with 1 line per protein and 1 column per condition will be to take the mean LogIntensities of the Runs per protein.

```{r get_runlevdata)}
tab_dp_msts_data <- dp_all_log2_noNorm$RunlevelData
```

#### Create wide tabular data  

```{r wide_tab_create}
tab_selection_msts_data <- dplyr::select(tab_dp_msts_data,
                                    Protein, LogIntensities, Patient = originalRUN, Group = GROUP_ORIGINAL)

tab_wide_msts_data <- pivot_wider(data = tab_selection_msts_data,
                                       names_from = Patient, values_from = LogIntensities) # make data wide

if(dir.exists(here::here("Tabular_data_Limma_input_format")) == FALSE){dir.create(here::here("Tabular_data_Limma_input_format"))}


write_csv(x = tab_wide_msts_data, path = here::here("Tabular_data_Limma_input_format/msstats_tabular_data_for_limma_input.csv"))
```




