---
title: "MQ-to-MSstats formating and normalization exploration (Sample report 1)"
author: "Miguel Cosenza"
date: "27 11 2019"
output: rmarkdown::github_document
---

```{r, echo = FALSE, include=FALSE}
knitr::opts_chunk$set(
  fig.path = "Figures_Sample_Report_1/",
  echo = TRUE
)
```
## Load required packages

```{r load_packages}
library(tidyverse)
library(MSstats)
```


## Load input data 

```{r load_input_data}
evidence <- read.table(file = here::here("Big_files/evidence_pancan_sc1.txt"),
                       sep = "\t",
                       header = TRUE)

annotation <- read.csv(file = here::here("Data/Pan_Can_Cohort/annotation_pancan_sc1.csv"),
                       header = TRUE)

proteingroups <- read.table(file = here::here("Data/Pan_Can_Cohort/proteinGroups_pancan_sc1.txt"),
                            sep = "\t",
                            header = TRUE)
```

## Transformation from Max Quant format to MSstats format 

Here I am evaluating different argument options for the `MaxQtoMSstatsFormat()` function in `MSstats`. 

### Set `removeProtein_with1Peptide = **FALSE**` (all default params)

This is the default parameter.

```{r mqtomstformat1}
if(dir.exists(here::here("MSstats_Exploration_Outputs")) == FALSE){dir.create(here::here("MSstats_Exploration_Outputs"))}

if(file.exists(x = here::here("MSstats_Exploration_Outputs/msts_data_w1pep.Rda")) == FALSE){ # check if Rda file with the data is already available. 
      
      msts_data_w1pep <- MaxQtoMSstatsFormat(evidence = evidence,
                                         annotation = annotation,
                                         proteinGroups = proteingroups,
                                         removeProtein_with1Peptide = FALSE)
      
      save(msts_data_w1pep, file = here::here("MSstats_Exploration_Outputs/msts_data_w1pep.Rda")) # save Rda file for faster loading later if required
      
      if(dir.exists(here::here("MSstats_Exploration_Outputs/MSstats_formated_tables")) == FALSE){dir.create(here::here("MSstats_Exploration_Outputs/MSstats_formated_tables"))}
      
      write.csv(x = msts_data_w1pep, file = here::here("MSstats_Exploration_Outputs/MSstats_formated_tables/msstats_data_protsw1pep.csv")) # protsw1pep means 'this file contains proteins that were identified only with 1 peptide". 
} else {
      
      load(here::here("MSstats_Exploration_Outputs/msts_data_w1pep.Rda")) # load the Rda file if available.
      
}
```

### Set `removeProtein_with1Peptide = **TRUE**` (all other params to default)

```{r mqtomstformat2}
if(dir.exists(here::here("MSstats_Exploration_Outputs")) == FALSE){dir.create(here::here("MSstats_Exploration_Outputs"))}

if(file.exists(x = here::here("MSstats_Exploration_Outputs/msts_data_noprotw1pep.Rda")) == FALSE){ # check if Rda file with the data is already available. 
      
      msts_data_noprotw1pep <- MaxQtoMSstatsFormat(evidence = evidence,
                                         annotation = annotation,
                                         proteinGroups = proteingroups,
                                         removeProtein_with1Peptide = TRUE)
      
      
      save(msts_data_noprotw1pep, file = here::here("MSstats_Exploration_Outputs/msts_data_noprotw1pep.Rda")) # save Rda file for faster loading later if required
      
      if(dir.exists(here::here("MSstats_Exploration_Outputs/MSstats_formated_tables")) == FALSE){dir.create(here::here("MSstats_Exploration_Outputs/MSstats_formated_tables"))}
      
      write.csv(x = msts_data_noprotw1pep, file = here::here("MSstats_Exploration_Outputs/MSstats_formated_tables/msstats_data_noprotw1pep.csv")) # noprotw1pe means 'this file does not contains proteins that were identified only with 1 peptide". 
} else {
      
      load(here::here("MSstats_Exploration_Outputs/msts_data_noprotw1pep.Rda")) # load the Rda file if available.
      
}
```

#### Check the difference in the number of proteins when using the different parameters 


**Number of proteins when including those identified with 1 peptide (default)**
```{r diff_n_prot_byformating_options1}
length(unique(msts_data_w1pep$ProteinName))
```

**Number of proteins when not including those identified with 1 peptide**
```{r diff_n_prot_byformating options2}
length(unique(msts_data_noprotw1pep$ProteinName))
```

```{r include=FALSE}
uniqpepdat1 <- tibble(N = c(length(unique(msts_data_w1pep$ProteinName)), length(unique(msts_data_noprotw1pep$ProteinName))),
                      param = c("w1pep","noprotw1pep"))
```

```{r barplot1}
barplot1 <- ggplot(uniqpepdat1,
                   aes(y = N,
                       x = param))+
            geom_col()
print(barplot1)
```

In this case I'll continue with normalization on the data that includes proteins identified with 1 peptide.

## Correction of protein IDs  

Get the first listed protein ID when many Uniprot IDs are listed into one Protein Group (identified peptides matching several proteins in the database)

```{r correct_proteinIDs}
msts_formated_data <- dplyr::mutate(msts_data_w1pep,
                             ProteinName = stringr::str_remove_all(ProteinName, ";.*$")) %>%
                      dplyr::mutate(ProteinName = stringr::str_trim(ProteinName))
```

## Testing MSstats `dataProcess()` options  

### Testing options for `normalization` (all features selected and log transformation = 2)  

#### Equalize Medians normalization  
```{r dataprocess1}
dp_all_log2_eqmNorm <- dataProcess(msts_formated_data,
                                   logTrans=2,
                                   normalization="equalizeMedians",
                                   nameStandards=NULL,
                                   address="",
                                   fillIncompleteRows=TRUE,
                                   featureSubset="all",
                                   remove_uninformative_feature_outlier=FALSE,
                                   n_top_feature=3,
                                   summaryMethod="TMP",
                                   equalFeatureVar=TRUE,
                                   censoredInt="NA",
                                   cutoffCensored="minFeature",
                                   MBimpute=TRUE,
                                   remove50missing=FALSE,
                                   maxQuantileforCensored=0.999,
                                   clusters=NULL)
```

#### Quantile normalization  
```{r dataprocess2}
dp_all_log2_quatNorm <- dataProcess(msts_formated_data,
                                   logTrans=2,
                                   normalization="quantile",
                                   nameStandards=NULL,
                                   address="",
                                   fillIncompleteRows=TRUE,
                                   featureSubset="all",
                                   remove_uninformative_feature_outlier=FALSE,
                                   n_top_feature=3,
                                   summaryMethod="TMP",
                                   equalFeatureVar=TRUE,
                                   censoredInt="NA",
                                   cutoffCensored="minFeature",
                                   MBimpute=TRUE,
                                   remove50missing=FALSE,
                                   maxQuantileforCensored=0.999,
                                   clusters=NULL)
```

### Testing options for `featureSubset` selection (log transformation = 2 and 'equalizeMedians' normalization)  

#### 
